---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
---
## Introduction
This is a document shows all the steps to complete the peer assessment 1
  
1.Set global default options, make the figure path for saving image files.
```{r setoptions, echo=TRUE}
require(knitr)
opts_chunk$set(echo = TRUE, fig.path = "figure/", fig.width = 6, fig.height = 6)
```
  
2.Load the library for processing
```{r loadlibraries}
library(dplyr)
library(lubridate)
library(lattice)
```
  
  
## I. Loading and preprocessing the data
1.Unzip and load the data
```{r getdata}
unzip("activity.zip", overwrite = TRUE)
allActivity <- read.csv("activity.csv",header = TRUE,sep=",",na.strings = "NA", stringsAsFactors = FALSE)
```
  
2.Transform the column "date" into date format "YY-mm-dd"
```{r formatdate}
allActivity$date <- as.Date(allActivity$date, "%Y-%m-%d")
```
  
  
## II. Calculate the average total number of steps taken per day
1.Group the data by date, and calculate the total number of steps taken per day. 
```{r groupbydate}
dailySteps <- allActivity %>%
        group_by(date) %>%
        summarize(sumStep = sum(steps))
```
  
2.Plot the histogram of the total number of steps taken each day.
```{r plothist}
hist(dailySteps$sumStep, col="green", 
     main = "Histogram of Total Number of steps taken per day",
     xlab = "Total number of steps",
     breaks = 20)
```
  
3.Calculate the mean and median of the total number of steps taken per day, ignore the missing values.
```{r calculatedailySteps}
avg_dailySteps <- mean(dailySteps$sumStep, na.rm = TRUE)
med_dailySteps <- median(dailySteps$sumStep, na.rm = TRUE)
```
### So the mean is `r avg_dailySteps`, the median is `r med_dailySteps`.
  
  
## III. Explore the average daily activity pattern
1.Group the data by interval, and calculate the averaged steps taken across all days group by interval, ignore the missing value.
```{r groupbyinterval}
intervalMean <- allActivity %>%
        group_by(interval) %>%
        summarize(avgStep = mean(steps, na.rm = TRUE))
```
  
2.Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.
```{r plotseries}
with(intervalMean, plot(interval,avgStep,
                        type = "l",
                        main = "Average steps taken per 5-min interval",
                        xlab = "5-minute interval",
                        ylab = "Average number of steps")
     )
```
  
3.Find the 5-minute interval, which contains the maximum number of steps, on average across all the days in the dataset.
```{r findmaxinterval}
intervalmax <- intervalMean %>%
        filter(avgStep == max(avgStep))
  
```
### On average across all the days in the dataset, the 5-minute interval: `r intervalmax$interval`, contains the maximum number of steps: `r intervalmax$avgStep`.
  

## IV. Imputing missing values
1.Calculate the total number of missing values in the dataset.
```{r countNA}
ctNArows <- allActivity %>%
        filter(is.na(steps)==TRUE) %>%
        nrow()
```
And the total number of missing values in the data is `r ctNArows`
  
2.Use the mean for each 5-minute interval to fill in all of the above missing values in the dataset.
```{r imputmissing}
Newsteps <- vector(length = nrow(allActivity))
for(i in 1:nrow(allActivity))
{
        if(is.na(allActivity$steps[i]))
        {
                Newsteps[i] <- filter(intervalMean, 
                             interval == allActivity$interval[i])$avgStep
        }
        else
        {
                Newsteps[i] <- allActivity$steps[i]
        }
}
```
  
3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r newdataset}
NewActivity <- allActivity
NewActivity$steps <- Newsteps
```
  
4.Make a histogram of the total number of steps taken each day with the new dataset with imputed missing data.
```{r calculatenewdataset}
dailySteps2 <- NewActivity %>%
        group_by(date) %>%
        summarize(sumStep = sum(steps))            
  
hist(dailySteps2$sumStep, col="red", 
     main = "Histogram of Total Number of steps taken per day",
     xlab = "Total number of steps",
     breaks = 20)
avg_dailySteps2 <- mean(dailySteps2$sumStep)
med_dailySteps2 <- median(dailySteps2$sumStep)
isImpactAvg <- ifelse(avg_dailySteps == avg_dailySteps2,
                   "the same as", "different from")
isImpactMed <- ifelse(med_dailySteps == med_dailySteps2,
                   "the same as", "different from")
isImpact <- ifelse(avg_dailySteps == avg_dailySteps2
                   & med_dailySteps == med_dailySteps2,
                   "no", "some")
```
### In the new data set, the mean is `r avg_dailySteps2`, is `r isImpactAvg` the original data set; while the median is `r med_dailySteps2`, is `r isImpactMed` the original.  
### So there is `r isImpact` impact of imputing missing data on the estimates of the total daily number of steps.
  
## V. Find the differences in activity patterns between weekdays and weekends.
1.Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r addweekday}
NewActivity <- NewActivity %>% 
        mutate(weekday = ifelse(wday(date)>1 & wday(date)<7,
                                "weekday","weekend"))
intervalMean2 <- NewActivity %>%
        group_by(weekday,interval) %>%
        summarize(avgStep = mean(steps)) 
```
  
2.Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r panelplot}
xyplot(avgStep ~ interval |weekday, 
       data=intervalMean2,type = "l", 
       layout = c(1,2), 
       main = "Between weekends and weekdays", 
       xlab = "Time of a day", ylab = "Number of steps")
```
  
